---
title: "WinePrediction_1"
author: "Mike Santoro"
date: "5/14/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(ggpubr)
library(boot)
```

## Introduction
The project presented below will explore the relationship between the outcome of two dependent variables of wine, value and quality, and their chemical composition: including measures such as acidity, residual sugar, and alcohol percentage. A predictive model for pricing and quality of wine can provide guidance to small wineries and vineyards, especially those from less established wine regions, so that they can more accurately price their inventory and ensure a better return on investment. It will explore different regression techniques and compare their effectiveness.

## The Data
We have taken a data-set from Kaggle: https://www.kaggle.com/uciml/red-wine-quality-cortez-et-al-2009 
This data-set inlcudes the measurements from the wine along with its quality rating.

```{r}
df<-read.table("data/winequality-red.csv",stringsAsFactors = FALSE,
                sep=",",header = TRUE)
head(df)
```

## Clean and Normalize the data. 
Our dataset already contains only predictive values and output, so we do not need to remove any descriptive columns.  We must normalize the values within the dataset to avoid any bias and remove the output variable (quality) since it's the prediction. 
```{r}

#Normalization
normalize <- function(x) {
return ((x - min(x)) / (max(x) - min(x))) }

wine.normal <- as.data.frame(lapply(df[,1:11], normalize))

head(wine.normal)

```



## Test Data-Frame and Working Data-Frame
Since our data-set our research question involves prediction will will randomly select a portion of data to use for overall effectiveness measurement. We plan to save about $5\%$ of the data for testing which ends up being $80$ values.

```{r}
set.seed(123)
dat.d <- sample(1:nrow(wine.normal),size=nrow(wine.normal)*0.8,replace = FALSE) #random selection of 90% data.
 
train.wine <- wine.normal[dat.d,] # 90% training data
test.wine <- wine.normal[-dat.d,] # remaining 10% test data

```

## Descriptive Statitics
We first will complete some initial plots to gain familiarity with the data. This will include boxplots, histograms, and qqplots.

### Histograms
```{r}
par(mfrow=c(2,2))
hist(train.wine$fixed.acidity)
hist(train.wine$volatile.acidity)
hist(train.wine$citric.acid)
hist(train.wine$residual.sugar)
```

```{r}
par(mfrow=c(2,2))
hist(train.wine$chlorides)
hist(train.wine$free.sulfur.dioxide)
hist(train.wine$total.sulfur.dioxide)
hist(train.wine$density)
```

```{r}
par(mfrow=c(2,2))
hist(train.wine$pH)
hist(train.wine$sulphates)
hist(train.wine$alcohol)
hist(train.wine$quality)
```

### Line Plots
```{r}
par(mfrow=c(2,2))
plot(train.wine$pH,df.data$quality)
plot(train.wine$sulphates, df.data$quality)
plot(train.wine$alcohol, df.data$quality)
plot(train.wine$quality, df.data$quality)
```


### Box Plots
```{r}
boxplot(train.wine$pH)
```



### QQPlots

```{r}
par(mfrow=c(2,2))
qqplot(train.wine$pH,train.wine$quality)
qqplot(train.wine$sulphates, train.wine$quality)
qqplot(train.wine$alcohol, train.wine$quality)
qqplot(train.wine$quality, train.wine$quality)
```





